Fast Fujisaki-Okamoto transformation using encrypt-then-mac and applications to Kyber

### 2024-08-19 etm-transform/manuscript/main.tex
things I need to work out

\section{Application to Kyber}

\subsection{Kyber with AE}
CRYSTALS-Kyber \cite{bos2018crystals}\cite{avanzi2019crystals} and ML-KEM \cite{key2023mechanism} are \texttt{IND-CCA2} secure key encapsulation mechanisms whose security depends on the hardness of the Module Learning with Error (MLWE) problem. For the construction of the \texttt{IND-CCA2} secure KEM, Kyber first constructs an \texttt{IND-CPA} public-key encryption scheme (which we will call \texttt{CPAPKE}), then applies the modular Fujisaki-Okamoto transformation \cite{hofheinz2017modular} to construct the key encapsulation mechanism (which we will call \texttt{CCAKEM}). Specifically, Kyber's round-3 submission uses the $U^{\not\bot}$ transformation, while ML-KEM uses the $U^{\not\bot}_m$ transformation. The routines of \texttt{CPAPKE} can be found in Algorithm 4, 5, 6 in \cite{avanzi2019crystals} and are largely identical between Kyber and ML-KEM.

@@

\subsection{MAC Performance}\label{sec:mac-performance}
We claim that the input $\mac$ only needs to be one-time existentially unforgeable. This is because besides the challenge ciphertext $(c^\ast, t^\ast)$, the adversary has no external resources from which it can obtain authenticated ciphertexts for which it does not know the decryption. Here we compare the performance of a variety of $\mac$ instantiations. Some of them are many-time secure while others are one-time secure. The standalone performance results are listed in table \ref{tbl:standalone-mac-perf}. For each choice of MAC, we checked the median (top) and average (bottom) CPU cycles (run on a 2019 MacBook Pro 16-inch) needed to sign 768, 1088, and 1568 bytes of data (respectively the ciphertext size for Kyber512, Kyber768, and Kyber1024).

>>>>

\section{Application to Kyber}
CRYSTALS-Kyber \cite{bos2018crystals}\cite{avanzi2019crystals} is an IND-CCA2 secure key encapsulation mechanism that first constructs an IND-CPA secure public key encryption scheme whose security is based on the Module Learning with Error Problem (MLWE), then applies a generic transformation using \emph{de-randomization} and \emph{re-encryption} \cite{hofheinz2017modular}. The use of \emph{re-encryption} for providing rigidity means that the decapsulation routine needs to run the encryption routine for verifying ciphertext integrity. Unfortunately for Kyber, the encryption routine consumes both more CPU cycles and carries a larger memory footprint. Therefore, applying an alternate transformation that replaces \emph{re-encryption} with a MAC verification will bring substantial performance enhancement. This is particularly applicable to communication protocols such as TLS 1.3, where clients, often constrained environments, need to run the decapsulation routine.

The \texttt{IND-CPA} PKE of Kyber (algorithms 4, 5, 6 in \cite{avanzi2019crystals}) is not \texttt{OW-PCA} secure. A plaintext-checking attack \cite{ravi2019generic} can recover the secret key using a few maliciously constructed plaintext-checking queries for each coefficient of a Kyber secret key (there are 512, 768, and 1024 coefficients in the secret key depending on the desired security level). However, we propose mitigation for plaintext-checking attack at protocol level by requiring each keypair to be used for decryption only once. If a decryption fails, then the secret key should be discarded and the key exchange terminated, and a new keypair should be generated to restart the key exchange. In fact, such ephemeral key exchange is already required in TLS 1.3 for forward secrecy, so requiring the one-time use of keypair does not introduce additional operational cost for protocols such as TLS 1.3.

@@

\subsection{Choosing message authentication code}\label{sec:mac-performance}
Given that each keypair is only used once, an adversary can only obtain one authenticated ciphertext, which means that the corresponding MAC key is also only used once per key exchange. This allows us to use MAC constructions that are only one-time secure for substantial computation efficiency gains. We compared the median and mean number of CPU cycles (Intel i7-9750H) used to sign Kyber's IND-CPA ciphertexts. The results are listed in \ref{tbl:standalone-mac-perf}.
